---
title: "Q3"
author: "YOUR NAME HERE"
format: html
---

## Exercise 3

Download the `numbats` data from [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-03-07/readme.md), following the instructions on the website.

a. Filter the data to be sightings after 2010.
b. Count the number of sightings for each day.
c. Make a stem-and-leaf of the sightings. Describe the shape. 
d. If you were to transform the data to symmetrise, what function would you recommend? Explain why or why not this would be recommended for this variable.
e. Following Tukey's EDA, identify which observation would be the median, Q1, Q3, and thus report the five-number-summary. Which statistic does Tukey refer to as the hinge?
f. Compute the step, inner fence, outer fence. Make a Tukey-style schematic plot, hand-drawn. Which observations would be considered to be outliers?
g. Use R to make a default boxplot. Which observations would be considered to be outliers? Why or why not is this a reasonable set of outliers? 


# Load the data

```{r}
library(tidyverse)
```

```{r}
numbats <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-03-07/numbats.csv')
```

# Answers

## Part A

```{r}
numbatsfilt <- numbats |>
  filter(!is.na(year)) |>
  filter(year > 2010)
```

## Part B

```{r}
numbats_daily_count <- numbatsfilt |>
  group_by(eventDate) |>
  summarize(daily_sightings = n())

numbats_daily_count
```

## Part C

```{r}
stem(numbats_daily_count$daily_sightings)
```

There is a very high skewness to the right and thus is asymmetrical. Moreover, it is non-uniform and less spread out as 1 is a huge mode (unimodel). This also shows less variability because the leaves are tightly packed towards the less number of sightings.

## Part D

```{r}
stem(sqrt(numbats_daily_count$daily_sightings))
stem(log(numbats_daily_count$daily_sightings))
```

There was an attempt to achieve symmetry where I tried using the square root and the log. However, there is still a fair amount of right skewness even after the transformations and thus, both methods were rejected. I did ponder upon using the box cox transformation but that requires no negative and 0 values, but our data has no sightings so that was also not feasible.

## Part E

The median is 2 sightings, the Q1 is 1 sighting, and the Q3 is 3.

## Part F

## Part G

```{r}
# Assuming numbats_daily_count contains the data
boxplot(numbats_daily_count$daily_sightings, main="R Default Box Plot")

```

